---
title: 'Lab week 8: binomial logistic regression'
author: "Sydney Mayes"
date: "2023-01-26"
output: html_document
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(jtools)
library(AICcmodavg)
```

# Pseudocode
* Examine our data (plots, tables, summary stats) - exploratory analysis
* Identify a question
* Wrangle the data
* Identify some candidate models
* Select among the candidate models using AIC/BIC
* Select among the candidate models using K-fold cross validation
* Select abont the candidate models using area under Receiver Operating Characteristic Curve

```{r}
GGally::ggpairs(penguins %>%  select(species, bill_length_mm:sex), 
                aes(color = species))
```
```{r}
class(penguins$species)
levels(penguins$species)

adelie_chinstrap <- penguins %>% 
  filter(species %in% c('Adelie', 'Chinstrap')) %>% 
  mutate(species = fct_drop(species)) %>% 
  select(-year) %>% 
  drop_na()
# use %in% instead of == with a vector -- it is more like the logical or, wont exclude a variable
  
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm))+
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~ species) # split into different graphs as a function of species; creates two facets, one for adelie, one for chinstrap

ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm)) + 
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~ species)
```

## Let's do some binary logistic regression!

```{r}
f1 <- species ~ body_mass_g + flipper_length_mm + sex #formula 1
ad_chin_blr1 <- glm(formula = f1, data = adelie_chinstrap,
                    family = 'binomial')
ad_chin_blr1
levels(adelie_chinstrap$species) # tells you the reference species (the first one?)
summary(ad_chin_blr1)

blr1_tidy <- tidy(ad_chin_blr1) # puts summary stats in a data frame
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = sex))
```
```{r}
blr1_fitted <- ad_chin_blr1 %>% 
  broom::augment(type.predict = 'response') # pulls data from model and augments it; makes a nice table. Fitted results will tell you the chance of being an adelie or chinstrap based on model

ggplot(data = blr1_fitted, aes(x = flipper_length_mm, y = .fitted)) +
  geom_point(aes(color = sex, shape = species)) +
  geom_smooth(aes(color = sex), se = FALSE) +
  labs(x = 'Flipper length(mm)', y = 'Probability of outcome(Chinstrap)')
```
## Predictions for new values with predict()
```{r}
ex1 <- predict(ad_chin_blr1,
                 data.frame(sex = 'female', 
                            body_mass_g = 3410,
                            flipper_length_mm = 192),
                 type = 'response' ) #if you dont tell it the type, it will default to log odds
ex1

new_df <- data.frame(
  sex = c('male', 'female', 'female'),
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)
ex2 <- predict(ad_chin_blr1, new_df, type = 'response')
ex2
```

# Create a new binary logisitc model
```{r}
f2 <- species ~ bill_length_mm + body_mass_g
ad_chin_blr2 <- glm(formula = f2, data = adelie_chinstrap, family = 'binomial')
ad_chin_blr2
summary(ad_chin_blr2)
blr2_tidy <- broom::tidy(ad_chin_blr2)
blr2_tidy
```


```{r}
ggplot(adelie_chinstrap, aes(x = bill_length_mm, y = body_mass_g)) + geom_point(aes(color = species))
```
## Model Selection
```{r}
AICcmodavg::aictab(list(ad_chin_blr1, ad_chin_blr2))
# Mod2 is much better
AICcmodavg::bictab(list(ad_chin_blr1, ad_chin_blr2))
# Similar results
```

10-fold cross validation
```{r}
set.seed(123) # pretty arbitrary number, but if you use the same number it should give the same outputs for everyone
n_folds <- 10
fold_vec <- rep(1:n_folds, length.out = nrow(adelie_chinstrap))

ad_chin_kfold <- adelie_chinstrap %>% 
  mutate(fold = sample(fold_vec, size = n(), replace = FALSE))
ad_chin_kfold

# check key for for loop code, but instead of a for loop we will now use a different package purrr
```

purrr::map()
```{r}
x_vec <- 1:10
thing <- purrr::map(.x = x_vec, ### a sequence (vector, list)
                    .f = sqrt) ### a function to use, will apply this function to every element in the sequence
thing
# this operates like a for loop, but is way more efficient.

my_funct <- function(x, y, z) {
  return((x - y) ^ z)
}
thing2 <- purrr::map(.x = x_vec, 
                     .f = my_funct,
                     y = 2, z = 3)
thing2
```

```{r}
pred_acc <- function(x, y) {
  accurate <- ifelse(x == y, 1, 0)
  
  return (mean(accurate, na.rm = TRUE))
}


calc_fold <- function(i, fold_df, f) {
  kfold_test <- fold_df %>% 
    filter(fold == i)
  kfold_train <- fold_df %>% 
    filter(fold != i)
  
  kfold_blr <- glm(f, data = kfold_train, family = 'binomial')
  
  kfold_pred <- kfold_test %>% 
    mutate(blr = predict(kfold_blr, kfold_test, type = 'response'),
           pred = ifelse(blr > 0.50, 'Chinstrap', 'Adelie')) ## 0.5 is the threshold
  
  kfold_accuracy <- kfold_pred %>% 
    summarize(blr_acc = pred_acc(species, pred))
  
  return(kfold_accuracy)
  
}
```


```{r}
results1_purrr_df <- purrr::map(.x = 1:n_folds, 
                                .f = calc_fold, 
                                fold_df = ad_chin_kfold,
                                f = f1) %>% 
  bind_rows() %>% 
  mutate(mdl = 'f1')
```

```{r}
results2_purrr_df <- purrr::map(.x = 1:n_folds, .f = calc_fold, 
                               fold_df = ad_chin_kfold,
                               f = f2) %>%
  bind_rows() %>%
  mutate(mdl = 'f2')

results_purrr_df <- bind_rows(results1_purrr_df, results2_purrr_df) %>%
  group_by(mdl) %>%
  summarize(mean_acc = mean(blr_acc))
results_purrr_df

```


# Tidymodels version
```{r}
# glm is default, but there are more options (e.g., brulee), don't technically need to set for glm

### define model type
blr_model <- logistic_reg() %>% 
  set_engine('glm') # could change to brulee, etc.

### basic regression
blr_tidyfit_f1 <- blr_model %>% 
  fit(f1, data = adelie_chinstrap)
blr_tidyfit_f2 <- blr_model %>% 
  fit(f2, data = adelie_chinstrap)

blr_tidyfit_f1
blr_tidyfit_f2

blr_tidyfit_f1 %>% 
  tidy() #creates a table with estimates
blr_tidyfit_f2 %>% 
  glance() #glances at info in a different way

```

## Tidy kfold cross validation
```{r}
set.seed(345)

tidy_folds <- vfold_cv(adelie_chinstrap, v = 10)

### use workflow to bundle a model and a formula
blr_tidyfit_wf1 <- workflow() %>% 
  add_model(blr_model) %>% 
  add_formula(f1)

blr_tidy_cv_f1 <- blr_tidyfit_wf1 %>% 
  fit_resamples(tidy_folds)

collect_metrics(blr_tidy_cv_f1)


blr_tidyfit_wf2 <- workflow() %>% 
  add_model(blr_model) %>% 
  add_formula(f2)

blr_tidy_cv_f2 <- blr_tidyfit_wf2 %>% 
  fit_resamples(tidy_folds)

collect_metrics(blr_tidy_cv_f2)

```

## Area under the curve
```{r}
blr_f1_pred <- adelie_chinstrap %>% 
  mutate(predict(blr_tidyfit_f1, .)) %>%
  mutate(predict(blr_tidyfit_f1, ., type = 'prob'))
  # . means operate on data set you provided

blr_f1_pred %>% 
  roc_curve(truth = species, .pred_Adelie) %>% 
  autoplot()

```

```{r}

```

